---
title: ''
output: html_document
---

# Reproducible {#reproduce}

We want our code to be reproducible so that: it can be used by others (both for collaboration and to allow effective review and accountability); it keeps working over time (protected from external changes); and so that it can be easily reused by others in their own projects (**AMEND** - delete this last one?).

There are a number of steps that we can take to ensure that our code is as reproducible as possible.

## Manage project dependencies {#projdep}

Your project will depend on an number of external factors, such as software or packages. These dependencies may mean that your project won't work on others' machines or may not work on your machine at a later date (e.g. as external packages are updated over time). To ensure that this doesn't become an issue for your project, you should use some kind of dependency management tool.

### Dependency management tools {-}

+--------------+--------------------------------------------------------+
| Language     | Tools                                  | 
+==============+========================================================+
| R            | Use [Conda][1], [Packrat][2] or [Renv][3]. [**AMEND** - add comment on which is preferable?] |
+--------------+--------------------------------------------------------+
| Python       | Use [Conda][1] |
+--------------+--------------------------------------------------------+
| Javascript   | Include third party library dependencies in the project as `.js` files |
+--------------+--------------------------------------------------------+

[1]:https://github.com/moj-analytical-services/coffee-and-coding-public/blob/master/2019-10-30%20Conda/conda.pdf
[2]:https://rstudio.github.io/packrat/
[3]:https://blog.rstudio.com/2019/11/06/renv-project-environments-for-r/


### Include a git hash {-#githash}

[**AMEND** - is this still relevant/current?]

If practical, the output of your code should include the git hash of the code that produced it. By doing so, the analysis should be
more reproducible, there is no ambiguity about the specific code that was used to generate it.

#### R {-}
You can access the git hash using either of the following code:
snippets.
```{r, results="hide"}
library(git2r)
repo <- repository(".")
print(repository_head(repo))
```

or

```{r, results="hide"}
print(system("git rev-parse --short HEAD", intern = TRUE))
```

#### Python {-}
You can access the git hash using the following code:

``` {python}
import subprocess
def get_git_revision_hash():
    return subprocess.check_output(['git', 'rev-parse', 'HEAD'])
def get_git_revision_short_hash():
    return subprocess.check_output(['git', 'rev-parse', '--short', 'HEAD'])
```

## Abstractions {-#functions}

Abstractions (e.g. functions, packages, modules etc.) should be used where possible. This makes code easier to understand, maintain and extend.

It will often be the case that will be a pre-existing package or module that contains the functions you need for your project. The best way to find these is usually through a quick google search and, if you're struggling to find what you're looking for, it's worth asking on the relevant slack channel (e.g. #r and #python in the [ASD workspace](asdslack.slack.com)

If you end up using a piece of code three times, it's probably worth turning it into a function and separating it out into a separate script. For more information on how to write functions, see [here][13]. As a general rule, functions should be less than about 50 lines long. 

All non trivial functions should be documented using the programming language’s accepted standard:  
* For Python follow [PEP8](https://www.python.org/dev/peps/pep-0008/) and particularly [PEP257](https://www.python.org/dev/peps/pep-0257/).  
* For R, use `roxygen2`.
If your functions are short and well documented, there is often little need for additional code comments. 
[**AMEND** - check links here?]

[13]:https://github.com/moj-analytical-services/writing_functions_in_r


## Format {#format}

If the output is a report, the write up should be fully reproducible, or as close as possible.  

* Avoid workflows that require manually copying and pasting results between documents.    
* For Python, consider using Jupyter notebooks. For R, use `rmarkdown`.  

## Optimize for change {#change}
* Don’t try to solve every conceivable problem up-front, instead focus on making your code easy to change when needed.
* Don’t prematurely optimize - choose clarity over performance, unless there is a serious performance issue that needs to be addressed.
* Change can come in several forms, including hardware - your code will eventually be run on a colleague's machine or a server somewhere.  Without over-complicating things, write your code with this in mind.  For example, use relative paths (e.g. `./file_in_the_project_directory.R` rather than `/Users/my_username/development/my_project/file_in_the_project_directory.R`)


